### 2013-02-15 doc/icml2013/evaluation.tex

\todo{Fill it up with juicy bits once we actually have them graphs.}

@@

We implemented the low rank recovery optimization routines for $M_2$ and
$M_3$ using a proximal subgradient descent algorithm\citationneeded. The
algorithm was initialized with the regularized least squares solution
and was run to convergence in each instance. Both regularization
parameters, $\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as
$\frac{1}{n}$.

@@

We generated synthetic data using covariates from the unit hyper cube
and non-linear responses.  To conform to the identifiability criteria
discussed in \sectionref{sec:algo}, we considered only features that
would guarantee this condition.

As an example, one set of features we considered in the uni-dimensional
setting were $\{1, x, x^4, x^7\}$. The data and the curves fit using
Spectral Experts, EM with random initializations and EM initialized with
the parameters recovered using spectral experts are shown in
\figureref{fig:curves}.

@@

\begin{table*}[t]
\caption{Parameter Recovery ($N = 500,000$).}
\label{tbl:parameter-recovery}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

  \begin{tabular}{ >{\centering}p{5cm}<{\centering} r c c c }
\hline
\abovespace\belowspace
Features  & $K$ & Spectral & EM & Spectral + EM \\
\hline
\abovespace
$\{1$, $ x_1$, $ x_1^4\}$ 
  & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
  & 3 & 1.87 $\pm$ 1.20 & {\bf 0.33 $\pm$ 0.96} & 0.35 $\pm$ 1.23 \\
$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
& 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & {\bf 1.51 $\pm$ 1.77} \\
$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
& 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & {\bf 7.31 $\pm$ 2.47} \\
$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
& 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & {\bf 0.51 $\pm$ 1.12} \\
$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & {\bf 1.42 $\pm$ 2.17} \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table*}

@@

There is a lot of variation across random instances; some are easy for
EM to find the global minima and others more difficult. Moreover, simply
increasing the number of samples did not necessarily improve EM
performance. In general, we found that while our spectral algorithm did
not recover parameters extremely well, it provided a good initialization
for EM.

\todo{Plot: the histogram over errors (for EM?) Some plots I have
doesn't really show much}.

@@

To evaluate how robust the algorithm was to model mis-specification, we
removed large contiguous sections from $x \in [-0.5,-0.25] \cup
[0.25,0.5]$. The lower half of \tableref{tbl:parameter-recovery} reports
recovery errors in this scenario.

\begin{figure*}[t]
  \centering
  \subfigure[Complete Data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3.pdf}
  }
    \hspace{-2em}
  \subfigure[Incomplete Data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
  }
  \caption{}
  \label{fig:vs-n}
\end{figure*}

Highlight

\todo{We see that the error of spectral is actually less than EM after a point.}

Point: tradeoff between statistical error and computational error;
not enough data, spectral is actually a lot worse
given enough data, spectral+EM is much better

\paragraph{Effect of noise}

\paragraph{Effect of separation between components}

>>>>

@@

We implemented the low rank recovery optimization routines for $M_2$ and
$M_3$ using a proximal subgradient descent
algorithm.%\cite{candes11phaselift,tomioka2010estimation}. (Note sure about this citation and not required?)
The algorithm
was initialized with the regularized least squares solution and was run
for 500 steps or till convergence. Both regularization parameters,
$\lambda_n^{(2)}$ and, $\lambda_n^{(3)}$ were annealed as $\frac{1}{n}$.

@@

We generated synthetic data using covariates from the $b$-dimensional
unit hyper cube $\{-1,1\}^b$ and considered non-linear featurizations of
this data that conformed to the identifiability criteria discussed in
\sectionref{sec:algo}. The $\beta$ were drawn from a standard Gaussian.

As an example, one set of features we considered in the uni-dimensional
setting were $\{1, x, x^4, x^7\}$. The data and the curves fit using
Spectral Experts, EM with random initializations and EM initialized with
the parameters recovered using spectral experts are shown in
\figureref{fig:curves}. We note that even on well separated data such as
this, EM converged to the global optima only 13\% of the time.

@@

\begin{table*}[t]
\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ ($N = 500,000$).}
\label{tbl:parameter-recovery}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

  \begin{tabular}{ r r r c c c }
\hline
\abovespace\belowspace
Variables ($b$) & Features ($D$) & Components ($K$) & Spectral & EM & Spectral + EM \\
\hline
\abovespace
%$\{1$, $ x_1$, $ x_1^4\}$ 
  1 & 4 & 2 & 1.52 $\pm$ 0.71 & 0.28 $\pm$ 0.82 & {\bf 0.13 $\pm$ 0.55} \\
%$\{1$, $ x_1$, $ x_2$, $ x_1^2 x_2^2\}$ 
  2 & 5 & 2 & 2.92 $\pm$ 1.10 & 0.00 $\pm$ 0.00 & {\bf 0.00 $\pm$ 0.00} \\
  2 & 5 & 3 & 3.76 $\pm$ 1.93 & 0.43 $\pm$ 1.07 & {\bf 0.23 $\pm$ 0.76} \\
%$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2 \}$ 
  2 & 6 & 2 & 4.52 $\pm$ 1.67 & 0.63 $\pm$ 1.29 & {\bf 0.18 $\pm$ 0.66} \\
\belowspace
  2 & 6 & 5 & 6.78 $\pm$ 2.18 & 2.29 $\pm$ 1.79 & {\bf 1.77 $\pm$ 1.89} \\

%  2 & 5 & 3 & 1.87 $\pm$ 1.20 & {\bf 0.33 $\pm$ 0.96} & 0.35 $\pm$ 1.23 \\
% 2 & 6 & 5 & 5.27 $\pm$ 2.32 & 1.80 $\pm$ 1.80 & {\bf 1.51 $\pm$ 1.77} \\
%$\{1$, $ x_1$, $ x_2$, $ x_1 x_2^3$, $ x_1^2 x_2^2$, $ x_1^3 x_2$, $ x_1^3 x_2^4$, $ x_1^4 x_2^3 \}$ 
% 2 & 8 & 7 & 8.42 $\pm$ 1.66 & 7.41 $\pm$ 2.99 & {\bf 7.31 $\pm$ 2.47} \\
%$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
% 3 & 10 & 3 & 3.78 $\pm$ 0.90 & 0.67 $\pm$ 1.49 & {\bf 0.51 $\pm$ 1.12} \\
%$\{1$, $ x_1$, $ x_2$, $ x_3$, $ x_1 x_2 x_3^2$, $ x_1 x_2^2 x_3$, $ x_1^2 x_2 x_3$, $ x_1^2 x_2^2$, $ x_1^2 x_3^2$, $ x_2^2 x_3^2$, $ x_3^2\}$
% 3 & 10  & 7 & 9.97 $\pm$ 3.22 & 1.43 $\pm$ 1.97 & {\bf 1.42 $\pm$ 2.17} \\

\hline

\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table*}

@@

One of the main reasons for the high variance is the variation across
random instances; some are easy for EM to find the global minima and
others more difficult. In general, while the spectral algorithm did not
recover parameters extremely well, it provided a good initialization for
EM.

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.50\textwidth]{figures/hist.pdf}
  \caption{Histogram over recovery errors between the three algorithms. $b = 1, d = 4, k = 3, N = 500,000$}
  \label{fig:hist}
\end{figure}

To study the stability of the solutions returned by the spectral method,
consider the histogram in \figureref{fig:hist}, which shows the recovery
errors of the algorithms over 170 attempts a dataset with $b = 1, d = 4,
k = 3$. Typically, the spectral algorithm returned a stable solution.
When these parameters were close enough to the true parameters, we found
that EM almost always converged to the global optima. Randomly
initialized EM only finds the true parameters a little over 10\% of the
time and shows considerably higher variance.

\begin{figure*}[tb]
  \centering
  \subfigure[Well Data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1833-decay.pdf}
  }
    \hspace{-2em}
  \subfigure[Misspecified Data]{
    \includegraphics[width=0.50\textwidth]{figures/vs-n/1-8-3-3-rm.pdf}
  }
  \caption{Recovey error vs $n$  ($b = 1, d = 5, K = 3, N = 500,000$).}
  \label{fig:vs-n}
\end{figure*}

In \figureref{fig:vs-n}, we show how the recovery error varies as we get
more data. Each data point shows the mean error over 10 attempts, with
error bars. We note that the recovery performance of EM does not
particularly improve; this is expected because of the problem of local
optima. The spectral algorithm's error decays slowly, and as it gets
closer to zero, EM initialized at the spectral parameters finds the true
parameters more often as well. This behaviour highlights the trade-off
between statistical and computational error.

@@

\begin{table*}[tb]
\caption{Parameter Recovery Error $\|\theta^* - \hat \theta\|_F$ with misspecified data ($N = 500,000$).}
\label{tbl:parameter-recovery-mis}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}

  \begin{tabular}{ r r r c c c }
\hline
\abovespace\belowspace
Variables ($b$) & Features ($d$) & Components ($K$) & Spectral & EM & Spectral + EM \\
\hline
\abovespace
 1 & 4 & 2 &  2.94 $\pm$ 1.52 & 0.29 $\pm$ 0.85 &  {\bf 0.22 $\pm$ 0.72} \\
\belowspace
 2 & 5 & 3 &  3.82 $\pm$ 2.00 & 0.44 $\pm$ 1.12 &  {\bf 0.35 $\pm$ 1.00} \\
 2 & 6 & 5 &  9.89 $\pm$ 4.46 & {\bf 2.53 $\pm$ 1.77} &  2.69 $\pm$ 1.83 \\
 2 & 8 & 7 & 23.07 $\pm$ 7.10 & 9.62 $\pm$ 1.03 &  {\bf 8.16 $\pm$ 2.31}  \\
\hline

\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table*}

To evaluate how robust the algorithm was to model mis-specification, we
removed large contiguous sections from $x \in [-0.5,-0.25] \cup
[0.25,0.5]$ and ran the algorithms again.
\tableref{tbl:parameter-recovery-mis} reports recovery errors in this
scenario. The error in the estimates grows larger for higher $d$.
